Using WordNet and Short-Term Memory for Contextual Disambiguation

### 2017-04-06 Draft_2.tex
Continued editing oof draft 2, title has now changed

        \Huge
        \textbf{Using WordNet and a Short-Term Memory model for Contextual Disambiguation}
        \vspace{2cm}

@@

        \Large
        Wordcount - 9191\\
        (Wordcount produced using wc - w)\\
        Supervisor - Dr Dimitar Kazakov\\
        Department of Computer Science\\
        University of York\\
        20 November 2016

@@

Semantic Analysis is a problem which continues to draw attention from researchers, due to the lack of a truly capable solution \cite{NLPHandbook}. Most would agree that, though a good computational solution is not available, the human brain performs impressively well when confronted with the same problem. Given the example:

@@

%------------ Memory Models
\subsection{Psycholinguistics}
\label{sec:Psycholinguistics}
Language understanding is a problem which, it can be stated, is solved by the human brain. From this satatement, it can be derived that a computational solution could be effectively built around knowledge of the processes at work in the brain. The process of language comprehension can be described using the working memory model \cite{MemoryBaddeleyEysenkAnderson}.

According to Baddely et al. \cite{MemoryBaddeleyEysenkAnderson} there exist multiple, special purpose, memory structures within two main categories, the Short-term Store and the Long-term Store.

\subsubsection{Long-term Store}
\label{LongTerm}
The long-term store (LTS) contains semi-permanent information. Within the LTS, there exist Explicit and Implicit memory structures. The contents of the Implicit memory describe skills and methods of doing things, whereas the Explicit memory contains factual information\cite{MemoryBaddeleyEysenkAnderson}. When considering these structures, it can be seen that Explicit memory is of greater interest in the context of NLP.

@@

\subsubsection{Short-term Store}
\label{ShortTerm}
The short-term store (STS) is a structure of limited capacity, used to store items for periods usually of no more than a few seconds \cite{MemoryBaddeleyEysenkAnderson}. In 1955, G. Miller, based upon previous experimental results, concluded that the size of the STS existed in the realm of 7$\pm$2 items of information \cite{SevenPlusMinusTwo}.

In 1971, R. Atkinson and R. Shiffrin proposed a model of the STS \cite{ControlProcessesSTMAtkinson}. In this model, the STS can both send information to, and draw information from the LTS. Inputs from the sensory registers (memory structures holding information relating to inputs from senses) are also sent to the STS. Atkinson and Shiffrin proposed that, over time, the activation of items in the STS decreased; they went on to theorise that items could be lost from the STS, only when a new, more highly activated item could take its place. To counter this loss of activation, the authors discussed the control process, rehearsal. This process makes use of repetition to increase the activation of items in memory, decreasing their chance of loss.

@@

According to the multiple-access model, when presented with an ambiguious word, initially all corresponding concepts are activated \cite{AccessingLexicalAmbiguities}. The most appropriate concept is then chosen using context and frequency.

Previously, we established that the process of disambiguation begins with the activation of all possible word concepts. The context-sensitive model deals with the use of context and frequency in the selection of the most appropriate of these \cite{PsychologyOfLanguage}. In cases where the context is strong, i.e. the correct concept can be chosen using it's surrounding context, the context is primarily relied upon for disambiguation. In the opposite case, i.e. when context gives little indication of which is the correct concept, the most frequenly used concept is used, assuming it fits with the available context.

%------------ WORDNET
\subsection{Wordnet}
\label{Wordnet}
In 1990, it was noted by G. Miller et al. that current attemps to organise the english lexicon, i.e. conventional dictionaries, offered few benefits when used in conjunction with computers \cite{WN1Introduction}. Wordnet was an effort to produce a dictonary, containing more information than a conventional dictionary, that could be useful for computational applications.

\label{Synsets}Central to the design of wordnet, is the idea of synsets \cite{WN1Introduction}. The authors began using the assumption that all concepts can be uniquely defined by their set of synonyms (words which share like meaning). In most cases, this assumption holds true, though, in cases where more detail is required, a "gloss" was added \cite{WN1Introduction}.

Wordnet builds upon models of the semantic memory, such as that discussed in the \hyperref[LongTerm]{Long-term store section} \cite{WN1Introduction}. The overall structure relies on four main semantic relations:

@@

It is common knowledge that words can fall into one of a number of categories, nouns, adjectives, verbs and adverbs. G. Miller et al. note that, due to the differences in the relations between words in these categories, each type has differs in the structure they produce and are therefore held in different files \cite{WN1Introduction}. The proceeding subsections will go into each of thesecategories in more detail.

\subsubsection{Nouns}
\label{Nouns}
G. Miller et al. note that a noun can be defined using only its immediate hypernym, and how it differs from its hypernyms other hyponyms \cite{WN2Nouns}. From this, it can be seen that hyponymy is perhaps the most important relation in the organisation of nouns. For this reason, nouns form a hierachical structure in wordnet.

@@

Other than synonymy, nouns have three other important features \cite{WN2Nouns}:
\begin{itemize}
	\item Attributes
	\begin{itemize}
		\item The attributes of a noun consist of adjectives which distinguish it from other hyponyms of its hypernym, for example \{huge, green, fluffy\}.
	\end{itemize}

@@

@@

	\item Relational Adjectives
	\begin{itemize}
		\item In the phrase "maternal instinct", it can be seen that the adjective is derived from a noun, in this case "mother"; this is the defining feature of relational adjectives \cite{WN3Adjectives}. In wordnet, relational adjectives are linked to their noun form, meaning they don't posess their own structure, instead falling into that described in the \hyperref[Nouns]{Noun section}.
	\end{itemize}

@@

\subsubsection{Verbs}
\label{Verbs}
In C. Fellbaums 1990 paper, "English Verbs as a Semantic Net", she discussed the lack of true synonymy across the verb category \cite{WN4Verbs}. This is an issue for Wordnet's designers, with their reliance on \hyperref[Synsets]{synsets}. The author goes on to describe the solution, periphrases, the use of verb phrases to give more meaning to a simple verb. In the paper, the example synset, \{swim, travel through water\} was given.

@@

In 2004, J. Sedding and D. Kazakov proposed an extension to the TF-IDF algorithm, using \hyperref[Wordnet]{Wordnet}. The aim of the paper was to include extra information provided by wordnet (synsets and hypernyms) in conjuncton with \hyperref[PoSTag]{PoS tagging}, to provide an output more suited to the categorisation of documents \cite{SeddingKazakov}. The authors ran multiple experiments, in order to compare different usage of additional information. unfortunately, it was found that their method was less effective than TF-IDF alone. As the authors stated, this is likely due to the lack of disambiguation, i.e. all synsets were used for each word, adding noise.

\subsubsection{Neural Networks}
\label{sec:NeuralNetwork}
In recent years the use of Neural Networks, more specifically Deep Neural Networks, in the solving of computational problems has dramatically increased. More recently, this technique has been applied to the problem of natural language processing. As discussed in the \hyperref[sec:NLP]{Natural Language Processing section}, it was shown that assigning varyng labels to words and phrases (i.e. categorisation) made up a large proportion of the processes, a job neural networks are wellsuited to, and often associated with \cite{NeuralNetworks}.

The first process requrired in the use of neural networks, is the conversion of raw text into a vector-base representation, giving a sentance as a set of vectors. The sentence is then passed into the neural network \cite{NeuralNetworks}. The network proposed by R.Collobert and J. Weston in 2008 contained multiple layers, each serving its own purpose \cite{NeuralNetworks}. Initially the network identifies word features, before using this data to calculate the probability of each synset of each word.

@@

The author noted that, highly activated synsets could remain in the short-term memory indefinitely. To counter this issue, and to remain in line with the memory models discussed in the \hyperref[ShortTerm]{Short-term store section}, the process of forgetting was added to the model, by multiplying the activation value by a number found by experimentation \cite{MattBurkePrevious}. Forgetting occurs over time, decreasing the activations of items in the Short-term memory and the Semantic memory, the latter by a greater degree.

@@

As mentioned previously, the values of some variables were found using experimentation. In these experiments, four variables were altered \cite{MattBurkePrevious}:
\begin{itemize}
	\item Disambiguation Method
	\begin{itemize}
		\item[] The author found that STM-first was marginally better (1\% difference)
	\end{itemize}
	\item Short-term memory size
	\begin{itemize}
		\item[] It was found that a STM size of 5 was optimal, with a 67\% accuracy.
	\end{itemize}
	\item The Attenuation value
	\begin{itemize}
		\item[] The author found that an attenuation value of between 0.7 and 0.9 gave the best result, with an accuracy of 67\%.
	\end{itemize}
	\item Forgetfulness
	\begin{itemize}
		\item[] It was found that a small amount of forgetfulness (multiply activation by 0.95) in conjunction with a small difference in forgetfulness between Semantic and Short-term memories (multiplied by 1.05), produced the best result \cite{MattBurkePrevious}.
	\end{itemize}
\end{itemize}

@@

@@

\subsection{Problem Definition}
\label{sec:ProbDef}
Semantic analysis is the process of extracting meaning from a sentence's tokens. Each individual token in a sentence has potentially many possible meanings; an effective tool should select the correct meaning, based upon each word's surrounding context. Throughout the \hyperref[sec:PrevWork]{previous work section} of the Literature Review, it is clear that Semantic Analysis is a problem which remains unsolved.

Semantic analysis has many applications, these include:
\begin{itemize}
	\item \textbf{Voice Recognition} has, over recent years, become an increasingly useful feature in consumer electronics. Conceptually, it can be seen that a system's ability to respond to voice input would be improved by a better semantic analysis.

	\item \textbf{The Aquisition of Information for knowledge based systems} could occur automatically,using an effective semantic analysis system in conjunction with a large corpura.
\end{itemize}
This problem can be seen to be exceptionally relevant to current technologies, making it a worthy subject of this research.

Previously, \hyperref[sec:LSA]{statistical methods} have been used to solve this problem. These methods, as stated \hyperref[sec:LSA]{previously} only consider frequency in their definition, and so are limited in their ability to handle corpura where a single word can occur multiple times throughout, with differing definitions accross these instances.

It is well known that the human brain's process of semantic analysis, is very effective. The brain's disambiguation process most likely uses the \hyperref[LongTerm]{memory structures proposed by Baddeley}\cite{MemoryBaddeleyEysenkAnderson}, containing information about relationships between word meanings. For example, given the following sentence:

the reader knows that "party"refers to a political party, likely, as discussed in the \hyperref[sec:DisambiguationModels]{Disambiguation Models section}, due to its pairing with the word "government".

Previous attempts at the problem of semantic analysis, don't necessarily correlate with our knowledge of the human brain. Given this knowledge, it could be possible to produce an equivalent computational model, which may be more effective than purely statistical methods. For this reason, this project will attempt to make use of a model of the brain's disambiguation processes to build a system, capable of semantic analysis.

\subsection{Project Aims}
\label{sec:ProjectAim}
The overall aim of this project is to build a system capable of finding the correct definitions of each word in a corpus. This problem is divisible into a set of smaller problems.

\begin{enumerate}
	\item \textbf{A computational model of memory structures} \newline
		This model must based upon the memory structures, as described by Baddeley \cite{MemoryBaddeleyEysenkAnderson}. An effective implementation should include both short and long term stores, as well as an episodic buffer. The LTS includes information regarding semantic links between words, the STS should contian information about the local context of a word, and the Episodic Buffer should contain information about the wider context of a word.

	\item \textbf{A system for reading in a corpus} \newline
	An input corpus must be processed in such a way that the context of a section of text (i.e. a small set of sentences) is represented by the STS, and the context of the corpus read so far is represented by the Episodic Buffer. This should be done using activation and forgettingin the STS, so as to comply with R. Atkinson and R. Shiffrin's model \cite{ControlProcessesSTMAtkinson}.

	\item \textbf{A computational model of disambiguation} \newline
	This model should make use of the both the local context of a word and knowledge of its frequency, so as to comply with \hyperref[sec:DisambiguationModels]{the Multiple Access Model} \cite{PsychologyOfLanguage}. The contextual disambiguation of a word should make use of the semantic links between its own, and other local words definitions. This would be most appropriately done using the most apparent relation, \hyperref[Hypernym]{hyponomy} \cite{WN2Nouns, WN4Verbs}.

The completed project will be evaluable according to the above critera, with a completely successful system meeting all of these.

\subsection{Project Organisation}
\label{sec:ProjectOrg}

\section{Design and Implementation}
\label{sec:Implementation}
The application was implemented using the Python programming language, chosen for the availability of the Natural Language Toolkit (NLTK) \cite{NLTK}. NLTK provides an interface for use with Wordnet and a number of useful corpura (collections of tagged raw text), to be used as test data.

\subsection{Corpus Analysis}
\label{sec:CorpusAnalysis}
Using the corpus reader built into NLTK, an entire document can be analaysed as a series of hierachical lists \cite{NLTK}:

\begin{itemize}
	\item \textbf{Document}, is a list of sentences,
	\item \textbf{Sentences}, each of which is a list of words.
\end{itemize}

The analyser traverses each of these lists, through a series of loops, processing them a sentance at a time.

@@

The function recursively traverses the tree of hypernyms an example of which is shown in Figure \ref{fig:hypernymTree}, activating each synset until, the model gives a result of less than zero. As the function traverses through each hypernym, the amount they are activated by decreases, favouring less general synsets.

@@

Upon activation, the synset in question's activation is increased by an amount, dependant upon the hypernym model used (discussed later in this section). The same synset can be activated multiple times within the same sentence, an example of which is shown in Figure \ref{fig:HypernymAct}. In this case, a similar model to \hyperref[sec:MattBurke]{that proposed by M. Burke} \cite{MattBurkePrevious} is used: 
\[H = \sum\limits_{i=1}^N S_i\] 
The total acivation increase is equal to the sum of all synset activations within the sentence.

It can be seen that, in Listing \ref{lst:HypernymActivation},there exists a function, hypernymModel. This function is responsible for reducing the activation increase of hypernyms as they become more general (closer to the top of the tree). This function should favour most heavily, less general hypernyms, i.e. those activated by the first few recursions of activateHypernyms.

@@

When the depth = 0 (i.e. the activation of the root synset) the model would require the calculation of log(0), which is undefined. For this reason, an offset of 0.001 was added, making the final model:
\[f(x) = -b\log(\frac{x}{a}+0.001)\]
The size of the offset needed to be small to reduce its impact on the overall model.

@@

Upon activation, the synset in question may enter the memory structures, if it posesses a high enough activation. Both an STS and Episodic Buffer have been implemented, using python classes for each. Interactions between the two memory structures are handled by a Memory Controller. As discussed in the \hyperref[Wordnet]{WordNet section}, WordNet provides a good model of the semantic memory \cite{WN1Introduction},described in the \hyperref[LongTerm]{Long-term Store section}. For this reason it has been used heavily.

\subsubsection{STS}
\label{sec:ImplementedSTS}

The STS (Stm in psuedocode) can, in a simplified sense, be described as a list containing memory items. These memory items have been implemented in python as an object with two attributes, synset (immutable) and activation (mutable), based upon the contents of the STS in the working memory model. These memory items can be replaced, if a synset with greater activation appears. The process of synset replacement is handled by the swapLowestItem function shown in Listing \ref{lst:swapLowestItem}.

@@

A notable feature of the STS is its limited capacity, as described by G. Miller in 1955 \cite{SevenPlusMinusTwo}. It is for this reason that the ability to swap memory items is required. In the case that the STS isn't full, the new memory item is simply added with no swap taking place.

\subsubsection{Episodic Buffer}
\label{sec:ImplementedEpisodicBuffer}

The purpose of the episodic buffer, is to maintain a list of synsets which have occurred in the STS previously. This allows future activations of synsets which have already occured to receive a boost when activated in the future. Interactions with the Episodic buffer only occcur during the activation of synsets, therefore, all interactions are handled by the Memory Controller.

\subsubsection{Semantic Memory}
\label{sec:ImplementedSemanticMemory}

The Semantic Memory is responsible for maintaining all the semantic knowledge of the system. It was decided that WordNet provided the best solution, with its in depth knowledge of synsets and their relations to eachother \cite{WN2Nouns, WN4Verbs}.

NLTK's built in WordNet reader \cite{NLTK} was used to interact with WordNet. Throughout the system, synsets are used. Each of these is a reference to a synset present in WordNet, allowing their semantic relationships (accessed through their methods) to be used.

>>>>

        \Huge
        \textbf{Using WordNet and a Short-Term Memory model for Semantic Analysis}
        \vspace{2cm}

@@

        \Large
        Wordcount - 10640\\
        \small
        (Wordcount produced using wc - w, \\
        this includes all the body of the report,\\
         and Appendix A)\\
        \large
        \vspace{0.8cm}
        Supervisor - Dr Dimitar Kazakov\\
        Department of Computer Science\\
        University of York\\
        20 November 2016

@@

Semantic analysis is a problem which continues to draw attention from researchers, due to the lack of a truly capable solution \cite{NLPHandbook}. If found, a reliable semantic analysis system would have many applications, a subset of these being:

\begin{itemize}
	\item \textbf{Voice Recognition} has, over recent years, become an increasingly useful feature in consumer electronics. Conceptually, it can be seen that a system's ability to respond to voice input would be improved by a better semantic analysis.

	\item \textbf{The Aquisition of Information for knowledge based systems} could occur automatically,using an effective semantic analysis system in conjunction with a large corpura.
\end{itemize}

Most would agree that, though a good computational solution is not available, the human brain performs impressively well when confronted with the same problem. Given the example:

@@

In the introduction, the basic motivations behind this project were laid out. This literature review aims to describe, and discuss the previous research, upon which this project is built.The following section will cover Psycholinguistics, previous work on semantic analysis, as well as some of the tools that are used throughout the implementation.

%------------ Memory Models
\subsection{Psycholinguistics}
\label{sec:Psycholinguistics}
The understanding of Natural language is a problem with which the human brain performs extremely well. From this satatement, it can be derived that a computational solution could be effectively built around knowledge of the processes at work in the brain. The process of language comprehension can be described, in part, using the working memory model\cite{MemoryBaddeleyEysenkAnderson}.

According to Baddely et al. \cite{MemoryBaddeleyEysenkAnderson} there exist multiple, special purpose, memory structures within two main categories, the Short-term Memory and the Long-term Memory. The working memory model describes how information can pass between the different structures, either originating from the Long-term Memory, or sensory inputs \cite{MemoryBaddeleyEysenkAnderson}.

\subsubsection{Long-term Memory}
\label{LongTerm}
The long-term memory (LTM) contains semi-permanent information. Within the LTM, there exist Explicit and Implicit memory structures. The contents of the Implicit memory describe skills and methods of doing things, whereas the Explicit memory contains factual information\cite{MemoryBaddeleyEysenkAnderson}. When considering these structures, it can be seen that Explicit memory is of greater interest in the context of NLP.

@@

\subsubsection{Short-term Memory}
\label{ShortTerm}
The short-term memory (STM) is a structure of limited capacity, used to store items for periods usually of no more than a few seconds \cite{MemoryBaddeleyEysenkAnderson}. In 1955, G. Miller, based upon previous experimental results, concluded that the size of the STM existed in the realm of 7$\pm$2 items of information \cite{SevenPlusMinusTwo}.

In 1971, R. Atkinson and R. Shiffrin proposed a model of the STM \cite{ControlProcessesSTMAtkinson}. In this model, the STM can both send information to, and draw information from the LTM. Inputs from the sensory registers (memory structures holding information relating to inputs from senses) are also sent to the STM. Atkinson and Shiffrin proposed that, over time, the activation of items in the STM decreased; they went on to theorise that items could only be lost from the STM when a new, more highly activated item could take its place. To counter this loss of activation, the authors discussed the control process, rehearsal. This process makes use of repetition to increase the activation of items in memory, decreasing their chance of loss.

\subsubsection{Episodic Buffer}
\label{sec:EpisodicBuffer}
When considering the reading of text, it was found that participants were able to recall significantly more than 7$\pm$2 items of information. In 2000, a structure linking the STM and LTM was proposed by Baddely \cite{BaddeleyEpisodicBuffer}. The Episodic Buffer makes use of sensory data to store information too old to exist in the STM, without committing it to the LTM. This often occurs through the use of imagery to represent a set of concepts \cite{MemoryBaddeleyEysenkAnderson}. When applied to natural language, this memory structure allows the representation of a greated context than the STM alone.

@@

According to the multiple-access model, when presented with an ambiguious word, initially all corresponding concepts are activated \cite{AccessingLexicalAmbiguities}. The most appropriate concept is then chosen using a balance of context and frequency.

The context-sensitive model extends the multiple-access model, by more appropriately using context and frequency in the selection of the most appropriate of concept \cite{PsychologyOfLanguage}. In cases where the context is strong, i.e. the correct concept can be chosen using it's surrounding context, the context is primarily relied upon for disambiguation. In the opposite case, i.e. when context gives little indication of which is the correct concept, the most frequently used concept is used, assuming it fits with the available context.

%------------ WORDNET
\subsection{Wordnet}
\label{Wordnet}
In order to build a successful system, within the time bounds given by this project, a good, preexisting, model of semantic memory is required. One such example of this is Wordnet \cite{WN1Introduction}.

In 1990, it was noted by G. Miller et al. that current attemps to organise the english lexicon, i.e. conventional dictionaries, offered few benefits when used in conjunction with computers \cite{WN1Introduction}. Wordnet was an effort to produce a dictonary, containing more information than a conventional dictionary, that could be useful for computational applications.

\label{Synsets}Central to the design of wordnet, is the idea of synsets \cite{WN1Introduction}. The authors began using the assumption that all word meanings can be uniquely defined by their set of synonyms (words which share like meaning). In most cases, this assumption holds true, though, in cases where more detail is required, a "gloss" was added \cite{WN1Introduction}.

Wordnet builds upon models of the semantic memory, such as that discussed in the \hyperref[LongTerm]{Long-term memory section} \cite{WN1Introduction}. The overall structure relies on four main semantic relations:

@@

It is common knowledge that words can fall into one of a number of categories, nouns, adjectives, verbs and adverbs. G. Miller et al. note that, due to the differences in the relations between words in these categories, each type has differs in the structure they imply, and are therefore held in different files \cite{WN1Introduction}. The proceeding subsections will go into each of these categories in more detail.

\subsubsection{Nouns}
\label{Nouns}
G. Miller et al. note that a noun can be defined using only its immediate hypernym, and how it differs from its hypernyms other hyponyms \cite{WN2Nouns}. From this, it can be seen that hyponymy is perhaps the most important relation in the organisation of nouns. For this reason, nouns form a strong hierachical structure in wordnet.

@@

Other than synonymy and hyponymy, nouns have three other important features \cite{WN2Nouns}:
\begin{itemize}
	\item Attributes
	\begin{itemize}
		\item The attributes of a noun consist of adjectives which distinguish it from other hyponyms of its hypernym, for example \{huge, green, fluffy\}.
	\end{itemize}

@@

Currently, all the described relations, excluding functions have been implemented into WordNet. Function words are a proposed feature, though adding this semantic link may prove difficult \cite{WN2Nouns}. Each noun has a potentially huge amount of functions, adding all of these may prove to be redundant, therefore it may be more useful for the most common (i.e. the most useful) to be included.

@@

	\item Relational Adjectives
	\begin{itemize}
		\item In the phrase "maternal instinct", it can be seen that the adjective is derived from a noun, in this case "mother"; this is the defining feature of relational adjectives \cite{WN3Adjectives}. In wordnet, relational adjectives are linked to their noun form, meaning they do not posess their own structure, instead falling into that described in the \hyperref[Nouns]{Noun section}.
	\end{itemize}

@@

\subsubsection{Verbs}
\label{Verbs}
In C. Fellbaums 1990 paper, "English Verbs as a Semantic Net", she discussed the lack of true synonymy across the verb category \cite{WN4Verbs}. This is an issue for Wordnet's designers, with their reliance on \hyperref[Synsets]{synsets}. The author goes on to describe the solution, periphrases, the use of verb phrases to give more meaning to a simple verb. In the paper, the example synset, \{swim, travel through water\} was given\cite{WN4Verbs}.

@@

In 2004, J. Sedding and D. Kazakov proposed an extension to the TF-IDF algorithm, using \hyperref[Wordnet]{Wordnet}. The aim of the paper was to include extra information provided by wordnet (synsets and hypernyms) in conjuncton with \hyperref[sec:IntroContext]{PoS tagging}, to provide an output more suited to the categorisation of documents \cite{SeddingKazakov}. The authors ran multiple experiments, in order to compare different usage of additional information. unfortunately, it was found that their method was less effective than TF-IDF alone. As the authors stated, this is likely due to the lack of disambiguation, i.e. all synsets were used for each word, adding noise.

\subsubsection{Neural Networks}
\label{sec:NeuralNetwork}
In recent years the use of Neural Networks, more specifically deep learning Neural Networks, in the solving of computational problems has dramatically increased. More recently, this technique has been applied to the problem of natural language processing. As discussed in the \hyperref[sec:NLP]{Natural Language Processing section}, it was shown that assigning varying labels to words and phrases (i.e. categorisation) made up a large proportion of the processes, a job neural networks are wellsuited to, and are often associated with \cite{NeuralNetworks}.

The first process requrired in the use of neural networks, is the conversion of raw text into a numerix, vector-based representation, capable of being processed. The sentence is then passed into the neural network \cite{NeuralNetworks}. The network proposed by R.Collobert and J. Weston in 2008 contained multiple layers, each serving its own purpose \cite{NeuralNetworks}. Initially the network identifies word features, before using this data to calculate the probability of each synset of each word.

@@

The author noted that, highly activated synsets could remain in the short-term memory indefinitely. To counter this issue, and to remain in line with the memory models discussed in the \hyperref[ShortTerm]{Short-term memory section}, the process of forgetting was added to the model, by multiplying the activation value by a number found by experimentation \cite{MattBurkePrevious}. Forgetting occurs over time, decreasing the activations of items in the Short-term memory and the Semantic memory, the latter by a greater degree.

@@

As mentioned previously, the values of some variables were found using experimentation. In these experiments, four variables were altered \cite{MattBurkePrevious}:
\begin{itemize}
	\item Disambiguation Method
	\begin{itemize}
		\item[] The author found that STM-first was marginally better (1\% difference)
	\end{itemize}
	\item Short-term memory size
	\begin{itemize}
		\item[] It was found that a STM size of 5 was optimal.
	\end{itemize}
	\item The Attenuation value
	\begin{itemize}
		\item[] The author found that an attenuation value of between 0.7 and 0.9 gave the best result.
	\end{itemize}
	\item Forgetfulness
	\begin{itemize}
		\item[] It was found that a small amount of forgetfulness (multiply activation by 0.95) in conjunction with a small difference in forgetfulness between Semantic and Short-term memories (multiplied by 1.05), produced the best result \cite{MattBurkePrevious}.
	\end{itemize}
\end{itemize}

@@

The memory structures use may also have contributed to the result. Each item in the LTM had an activation \cite{MattBurkePrevious}. This activation is not present in the models developed by Baddeley, Shiffrin and Atkinson \cite{MemoryBaddeleyEysenkAnderson, ControlProcessesSTMAtkinson}.

\subsection{SemCor}
\label{sec:SEMCOR}
\cite{SEMCOR}
In order to test the system developed in this project, an input corpus is required. SemCor (semantically-tagged corpus) was found to be a useful corpus, for a number of reasons.

\begin{itemize}
	\item \textbf{Range of Topics} \\
	SemCor is a subset of the Brown corpus, taking documents from the large range of categories available \cite{SEMCOR}. The result is a varied corpus which will give a more representative view of performance than a smaller collection of documents.

	\item \textbf{PoS Tags} \\
	The corpus has already been syntactically parsed, meaning PoS tags already exist for all words in the corpus \cite{SEMCOR}. This allows the focus of the project to remain on semantic analysis, without the need to build extra sub-systems for preprocessing.

		\item \textbf{Ease of Reading} \\
	A reader for SemCor has already been implemented as part of the Natural Language Toolkit for python (NLTK) \cite{NLTK}. The ease of reading provided by NLTK further reduces the scale of the implementation.

	\item \textbf{Semantic tags} \\
	The most useful feature of SemCor are the semantic tags. These semantic tags provide a means for testing the accuracy of the system, by comparing its output to these tags. Each of these tags is a WordNet synset, the same format used as the output of the system, easing comparison \cite{SEMCOR}.
\end{itemize}

@@

The literature review has given a great deal of indication to the direction of the problem. Given these findings, the problem can be defined with a greater deal of precision. The following can be read as a description of how the implementation should be approached, as well as acting as a set of aims, against which the project will be evaluable.

The overall aim of this project is to build a system capable of finding the correct meaning for each word in an input corpus. Earlier, we established that the brain can reliably solve this problem and, though researchers are currently working to find a suitable computational solution, there exist successful psychological models concerning this process.

The problem described above can be divided into three separate sub-problems.

\subsection{Memory Structures}
\label{sec:PAMem}

The brain’s ability to conduct semantic analysis is reliant upon the presence of multiple special purpose memory structures, as described by the Working Memory Model. The most prominent of these are the short-term and long-term memories, both of which should be implemented by an effective system.  As part of the working memory model used by this system, information must pass between the each of these memory structures.

An implementation of an STM should have a finite size, falling in the range defined by G. Miller, 7$\pm$2.  Atkinson and Shiffrin’s model requires that the implemented STM contains items originating from a combination of the systems input, and LTM, each with an activation. The structure’s contents must also change, according the arrival of new items, so as to correctly represent the local context of a section of text.

The systems LTM (Semantic Memory) requires knowledge of a large number of word meanings. Each of the meanings should be coupled with useful information consisting of a number of semantic links, connecting them to other meanings. As discussed in section ??, WordNet is a capable, pre-existing database, which follows a similar structure to the LTM defined by Baddeley. For this reason, it should be used to ensure that as much semantic information can be contained in the LTS as possible within the time scale of the project.

The addition of an Episodic Buffer may also be beneficial. Baddeley’s proposal may prove difficult to implement fully within the scope of this project, so design decisions must be made to simplify the model. With that said, its functionality must remain reasonably consistent, the representation of wider context, for it to prove useful to the system.

\subsection{Reading System}
\label{sec:PARead}

In order for the system to effectively use the memory structures, a reading subsystem is required. In section ??, the input corpus, SemCor, was described. The reading system must be able to accept  this corpus, and feed it into the memory structures described previously. With the focus of this project being solely on semantic analysis, a reading system must also be able to read, and use all relevant information already existing in the input corpus (for example, PoS tags).

In the multiple access model, all relevant concepts are activated when a word is encountered.  A successful reading system should change the contents of the STM using a series of activations. If a concept already exists in the STM, Atkinson and Shiffrin’s model’s process of rehearsal should be employed, increasing the concepts actviation.

To ensure a good representation of local context in the STM, its contents should consist of relatively general concepts.  For this to occur, activation should make use of hyponomy. M. Burke’s findings on hypernym activation imply that a model will need to be developed, in order to decrease the activations as the hypernym tree is traversed.

\subsection{Disambiguation System}
\label{sec:PADisambiguate}

The multiple access model, and, by extension, the context-sensitive model, requires the usage of both context and frequency to disambiguate tokens.  A system for both methods will need to be developed to ensure that word meanings can be found, regardless of strength of context. There must exist some method of deciding when contextual disambiguation is appropriate, and frequency based methods are not (and vice versa).

A contextual based system must make use of the context held within the system by accessing the contents of the STM.  M. Burke found using a hyponymy to detect similarity between concepts to be reasonably effective. By using at least this semantic relation, a successful system will be able to select from a number of meanings, given the local context alone.

In cases where context is weak, the context-sensitive model tells us that word frequency is used. Given a word, there must exist some method of finding which of its meanings is most likely.

\section{Design and Implementation}
\label{sec:Implementation}
Given the findings of the literature review, and the definition of the problem given previously, an implementation can be designed. The following section describes the system implemented throughout this project, using psuedocode and diagrams to illustrate. The structure of the section roughly follows the processes conducted by the application, in order, when presented with an input corpus.

The application was implemented using the Python programming language, chosen for the availability of the Natural Language Toolkit (NLTK) \cite{NLTK}. NLTK provides an interface for use with Wordnet and a number of useful corpura, specifially SemCor, to be used as test data.

\subsection{System Input}
\label{sec:CorpusAnalysis}
The input of the system is a preprocessed document, structured as a series of of hierachical lists \cite{NLTK}:

\begin{itemize}
	\item \textbf{Document}, is a list of paragraphs.
	\item \textbf{Paragraphs}, each of which is a list of sentences.
	\item \textbf{Sentences}, each of which is a list of words.
\end{itemize}

Before a document can be inputted, it must have been syntactically analysed. Each word in a correctly preprocessed document will have a PoS tag, signifying its purpose in the sentance. This tag can the be used to minimise the size of the set of possible definitions.

\subsection{System Output}
\label{sec:SysOutput}
Given an input in the form descrbed \hyperref[sec:CorpusAnalysis]{previously}, after processing, the system will output a document tagged using WordNet synsets. Each synset in the output represents the individual meanings of each word in the document.

@@

The function recursively traverses the tree of hypernyms, an example of which is shown in Figure \ref{fig:hypernymTree}, activating each synset until, the model gives a result of less than zero. As the function traverses through each hypernym, the amount they are activated by decreases, favouring less general synsets.

@@

Upon activation, the synset in question's activation is increased by an amount, dependant upon the hypernym model used (discussed later in this section). The same synset can be activated multiple times within the same sentence, an example of which is shown in Figure \ref{fig:HypernymAct}. In this case, a similar model to \hyperref[sec:MattBurke]{that proposed by M. Burke} \cite{MattBurkePrevious} is used: 
\[H = \sum\limits_{i=1}^N S_i\] 
The total acivation increase of a synset is equal to the sum of all its activations within the sentence.

It can be seen that, in Listing \ref{lst:HypernymActivation},  there exists a function, hypernymModel. This function is responsible for reducing the activation increase of hypernyms as they become more general (closer to the top of the tree). This function should favour most heavily, less general hypernyms, i.e. those activated by the first few recursions of activateHypernyms. This prevents more general, and therefore less contextually relevant, synsets from dominating the STM.

@@

When the depth = 0 (i.e. the activation of the root synset) the model would require the calculation of log(0), which is undefined. For this reason, an offset was added, with a value of 0.001 being used to reduce its impact on the output of the function, making the final model:
\[f(x) = -b\log(\frac{x}{a}+0.001)\]

@@

Upon activation, the synset in question may enter the memory structures, if it posesses a high enough activation. Both an STM and Episodic Buffer have been implemented, using python classes for each. Interactions between the two memory structures are handled by a Memory Controller.

\subsubsection{STM}
\label{sec:ImplementedSTM}

The STM (Stm in psuedocode) can, in a simplified sense, be described as a list containing memory items. These memory items have been implemented in python as an object with two attributes, synset (immutable) and activation (mutable), based upon the contents of the STM in the working memory model. These memory items can be replaced, if a synset with greater activation appears. The process of synset replacement is handled by the swapLowestItem function shown in Listing \ref{lst:swapLowestItem}.

@@

A notable feature of the STM is its limited capacity, as described by G. Miller in 1955 \cite{SevenPlusMinusTwo}. It is for this reason that the ability to swap memory items is required. In the case that the STM is not full, the new memory item is simply added with no swap taking place.

In this implementation, there exist two separate STMs for verbs and nouns. Due to the limited information regarding noun-verb relationships currently existing in WordNet, combining nouns and verbs in the same STM offers no benefit to the system, as nouns and verbs do not contribute to each other's activations.

\subsubsection{Episodic Buffer}
\label{sec:ImplementedEpisodicBuffer}
The Episodic Buffer described in section \ref{sec:EpisodicBuffer} makes use of imagery generated by sensory data to store information \cite{BaddeleyEpisodicBuffer}. A memory structure modelling this process exactly would not be possible within the scope of this project, therefore, a simplified model had to be created. In this simplified model, the Episodic Buffer still provides a memory for wider context, but does so using a synsets, not imagery.

The purpose of the episodic buffer, is to maintain a list of synsets which have occurred in the STM previously. This allows future activations of synsets which have already occured to receive a boost when activated in the future. Interactions with the Episodic buffer only occcur during the activation of synsets, therefore, all interactions are handled by the Memory Controller.

\subsubsection{Semantic Memory}
\label{sec:ImplementedSemanticMemory}
The Semantic Memory is responsible for maintaining all the semantic knowledge of the system. It was decided that WordNet provided the best solution, with its in depth knowledge of synsets and their relations to one another \cite{WN2Nouns, WN4Verbs}.
